---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
``` {r load_data}
unzip ("activity.zip")
rawdata <- read.csv("activity.csv")
```

```{r load_library, echo=FALSE, message=FALSE }
library(dplyr)
library(lattice)
```

## What is mean total number of steps taken per day?

### 1. Total number of steps taken per day
``` {r mean_steps}
steps_per_day_gp <- group_by(rawdata, date)
steps_per_day <- data.frame(summarize(steps_per_day_gp, dailysteps = sum(steps, na.rm = T)))
steps_per_day
```

### 2. Histogram of total number of steps taken per day
``` {r}
hist(steps_per_day$dailysteps)
```

### 3. Mean and median of the total number of steps taken per day
```{r}
meansteps <- format(mean(steps_per_day$dailysteps),big.mark=",",scientific=FALSE)
mediansteps<- format(median(steps_per_day$dailysteps),big.mark=",",scientific=FALSE)
```

The mean and median of the total number of steps taken per day are `r meansteps` and `r mediansteps` respectively

## What is the average daily activity pattern?
```{r}
steps_per_interval_grp <- group_by(rawdata, interval)
steps_per_interval <- data.frame(summarize(steps_per_interval_grp, averagesteps = mean(steps, na.rm =T)))
plot(steps_per_interval$interval,steps_per_interval$averagesteps, type ='l', xlab="5-minute interval", ylab="Average number of steps")
maxsteps <- steps_per_interval[which.max(steps_per_interval$averagesteps),1]
```

On average across all the days in the dataset, interval `r maxsteps` contains the maximum number of steps.

## Imputing missing values
```{r}
missingvalues<-sum(is.na(rawdata[1]))
```
There are `r missingvalues` rows with NAs.

```{r}

steps_per_day_gp_imputed <- mutate(steps_per_day_gp, dailystepsmean = mean(steps, na.rm = T))
steps_per_day_gp_imputed$dailystepsmean[is.na(steps_per_day_gp_imputed$dailystepsmean)] <-0
steps_per_day_gp_imputed <- mutate(steps_per_day_gp_imputed, steps2 = steps)
steps_per_day_gp_imputed$dailystepsmean[is.na(steps_per_day_gp_imputed$dailystepsmean)] <-0
steps_per_day_gp_imputed <- mutate(steps_per_day_gp_imputed, steps2 = steps)
steps_per_day_gp_imputed$steps2[is.na(steps_per_day_gp_imputed$steps2)] <- steps_per_day_gp_imputed$dailystepsmean[is.na(steps_per_day_gp_imputed$steps2)]

steps_per_day_gp_imputed <- group_by(steps_per_day_gp_imputed, date)
steps_per_day_imputed <- data.frame(summarize(steps_per_day_gp_imputed, dailysteps = sum(steps2, na.rm = T)))


hist(steps_per_day_imputed$dailysteps)

```

```{r}
meanstepsimputed <- format(mean(steps_per_day_imputed$dailysteps),big.mark=",",scientific=FALSE)
medianstepsimputed <- format(median(steps_per_day_imputed$dailysteps),big.mark=",",scientific=FALSE)
```
The mean and median of the total number of steps taken per day using imputed data are `r meanstepsimputed` and `r medianstepsimputed` respectively.

The values are exactly the SAME as the estimates computed above.

For this set of data, there is no impact imputing missing data.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
steps_per_interval_grp$daytype <- ifelse(weekdays(as.Date(steps_per_interval_grp$date)) %in% c("Saturday", "Sunday"), "weekend", "weekday")
steps_per_interval_grp <- group_by(steps_per_interval_grp, interval, daytype)
steps_per_interval <- data.frame(summarize(steps_per_interval_grp, averagesteps = mean(steps, na.rm =T)))

xyplot(averagesteps ~ interval | daytype, data = steps_per_interval, type="l", cex = 1.5, xlab="Interval", ylab="Number of steps", layout=c(1,2))


```